---
title: "Essie"
output: html_document
---
```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = 
                "D:\\BOULOT\\Projects\\Xmas\\Abbotts\\DATA\\CLEANING\\ABBOTTS")

```

#We assume that the data have been preprocessed to produce a suitable
#set of twilight data and a consistent initial path.

#Load `SGAT` and `BAStag`
```{r}
library(SGAT)
library(raster)
source("SGAT/colours.r")
```

## Configuration

Load basic configuration info.  We don't actually need a detailed
initial path, `x0` is only used for the fixed points.
```{r}
source("GLS_files/4063/4063_2_config.r")
if(diff(range(xlim))>360) xlim <- c(0,360)
twl <- read.csv(paste0("GLS_files/", tag, "/", tagdata, "_twilight.csv"),header=T)
twl$Twilight <- as.POSIXct(twl$Twilight,"GMT")
path <- read.csv(paste0("GLS_files/", tag, "/", tagdata, "_path.csv"), header=T)
path <- list(time=as.POSIXct(path$time,"GMT"),x=cbind(path$lon,path$lat))
fixed <- ifelse(twl$Marker <= nrow(fixed.locations),twl$Marker,0)
```

## Model Parameters

Set the model parameters controlling the speed distribution and the
distribution of twilight errors
```{r}
threshold
zenith
zenith0
alpha
beta
```

Create the raster that defines the grid
```{r}
library(maptools)
data("wrld_simpl")
## Cells per degree
n <- 1
grid <- raster(nrows=n*diff(ylim),ncols=n*diff(xlim),
               xmn=xlim[1],xmx=xlim[2],ymn=ylim[1],ymx=ylim[2],
               crs=proj4string(wrld_simpl))
grid <- cover(rasterize(elide(wrld_simpl,shift=c(360,0)),grid,1,silent=TRUE),
              rasterize(wrld_simpl,grid,1,silent=TRUE))
## Mask must be non-zero for sea
grid <- is.na(grid)
plot(grid)
```

## SST

Incorporate a contribution to the likelihood that assumes
the observed sea surface temperature is Normally distributed about
remotely sensed SST.
```{r}
sst <- stack("SGAT/sst.wkmean.1990-present.nc",varname="sst",quick=TRUE)
indices <- .bincode(twl$Twilight,as.POSIXct(strptime(getZ(sst),"%Y-%m-%d","GMT"),
                                            "GMT"),right=FALSE)
logp0 <- function(k,x) {
  x[,1] <- x[,1]%%360
  if(is.na(twl$SST[k])) {
    0
  } else {
    dnorm(twl$SST[k],extract(sst[[indices[k]]],x),2,log=T)
  }
}
```

## Model Fit

We extract the initial path x0 and the fixed locations
```{r}
fixed <- fixed > 0
x0 <- as.matrix(path$x)
```

Define a model
```{r}
model <- essieThresholdModel(twl$Twilight,twl$Rise, twilight.model = "LogNormal",
                               alpha=alpha,beta=beta,
                               logp0=logp0,
                               x0=x0,fixed=fixed,zenith=zenith0,)
```

Fit the model
```{r}
fit <- essie(model,grid,epsilon1=1.0E-4)
```

## Summary

Compare the final maximum a posteriori (map) estimate with the map
estimate from just the forward and backward passes.  If these were not
in loose agreement, we would be concerned about the fitted path.
```{r,fig.show='hold'}
opar <- par(mfrow=c(2,2))
plot(fit$grid,legend=FALSE,col=c(map1.col,"white"),main="Track")
lines(essieMode(fit)$x,type="l",col=trk.col)
plot(fit$grid,legend=FALSE,col=c(map1.col,"white"),main="Forward")
lines(essieMode(fit,type="forward")$x,type="l",col=trk.col)
plot(fit$grid,legend=FALSE,col=c(map1.col,"white"),main="Backward")
lines(essieMode(fit,type="backward")$x,type="l",col=trk.col)
par(opar)
```

Compare the final posterior mean estimate with the estimate from just
the forward and backward passes.  Again, if these were not in loose
agreement, we would be concerned about the fitted path.
```{r,fig.show='hold'}
opar <- par(mfrow=c(2,2))
plot(fit$grid,legend=FALSE,col=c(map1.col,"white"),main="Track")
lines(essieMean(fit)$x,type="l",col=trk.col)
plot(fit$grid,legend=FALSE,col=c(map1.col,"white"),main="Forward")
lines(essieMean(fit,type="forward")$x,type="l",col=trk.col)
plot(fit$grid,legend=FALSE,col=c(map1.col,"white"),main="Backward")
lines(essieMean(fit,type="backward")$x,type="l",col=trk.col)
par(opar)
```

Plot the posterior for a selection of times
```{r,fig.show='hold'}
opar <- par(mfrow=c(2,2),mar=c(3,3,4,1))
ks <- round(seq(1,length(model$time),length=9)[c(2,4,6,8)])
for(k in ks) {
  plot(fit$grid,legend=FALSE,col=c(map1.col,"white"),main=fit$time[k])
  plot(essieRaster(fit,k),add=T)
}
par(opar)
```

Write the estimated track (the sequence of posterior modes) adjusted
for fixed locations
```{r}
path <- essieMean(fit)
path$x[fixed > 0,] <- x0[fixed > 0,]
write.csv(cbind.data.frame(time=path$time,path$x),
          paste0("GLS_files/", tag, "/", tagdata,"_essieSST.csv"),row.names=FALSE)
```

<!--
```{r VersionInfo,results='asis',echo=FALSE}
cat(R.version.string,"\n",
    "SGAT: ",as.character(packageVersion("SGAT")),"\n",
    "BAStag: ",as.character(packageVersion("BAStag")),sep="")
```
-->
