---
title: "Estelle SST"
output: html_document
---


We assume that the data have been preprocessed to produce a suitable
set of twilight data and a consistent initial path.

Load `SGAT` and `BAStag`
```{r}
require(knitr)
opts_knit$set(root.dir = "C:/Users/2029406/Desktop/Jo_ABBOTTS/ABBOTTS")
library(SGAT)
library(BAStag)
library(tictoc)
source("SGAT/colours.r")
```


## Configuration

Load the configuration data generated in the preprocessing stage
```{r}
tic("Running time")
source("GLS_files/0585/0585_config.r")
if(diff(range(xlim))>360) xlim <- c(0,360)
twl <- read.csv(paste0("GLS_files/", tag, "/", tagdata, "_twilight.csv"),header=T)
twl$Twilight <- as.POSIXct(twl$Twilight,"GMT")
path <- read.csv(paste0("GLS_files/", tag, "/", tagdata, "_essieSST.csv"),header=T)
path <- list(time=as.POSIXct(path$time,"GMT"),x=cbind(path$lon,path$lat))
fixed <- ifelse(twl$Marker <= nrow(fixed.locations),twl$Marker,0)
```


## Model Parameters

Set the model parameters controlling the speed distribution and the
distribution of twilight errors
```{r}
threshold
zenith
zenith0
alpha
beta
```


## Prior + SST

We construct a sea mask that indicates when points on the path are at
sea or fall out of the bounding box defined by `xlim` and `ylim` and use
this to define a log prior that is applied to both the x and z
points, and incorporate a contribution to the likelihood that assumes
the observed sea surface temperature is Normally distributed about
the Reynolds optimally interpolated SST.
```{r}
source("SGAT/SST.r")
## Load sst data
reynolds <- sstLogp(twl,xlim=xlim,ylim=ylim,land=F,sd=2,
              src="Reynolds",path=
                "C:/Users/2029406/Desktop/Jo_ABBOTTS/ABBOTTS/SGAT")

# Define land mask
is.sea <- landMask(xlim=xlim,ylim=ylim,n=4,land=F)

## Define the log prior for x and z
log.prior.x <- function(p)  {
  logp <- reynolds$logp(p)
  f <- is.sea(p)
  ifelse(!is.na(f) & f,logp,-1000)
}
log.prior.z <- function(p)  {
  f <- is.sea(p)
  ifelse(!is.na(f) & f,0,-1000)
}
```

## Initialization run

Ideally the initial path would be chosen to meet the constraints imposed
by the data, but this is no always easy, and so an initialization run
is conducted.


We derive the initial paths x0 and z0 and the fixed locations from the edited path.
```{r}
x0 <- as.matrix(path$x)
z0 <- trackMidpts(x0)
fixedx <- fixed > 0
```

When the model is defined, the `logp.x` and `logp.z` arguments are
used to incorporate contributions from the prior to implement the sea
mask

```{r}
model <- thresholdModel(twl$Twilight,twl$Rise,
                         twilight.model="ModifiedLogNormal",
                         alpha=alpha,beta=beta,
                         logp.x=log.prior.x,logp.z=log.prior.z,
                         x0=x0,z0=z0,zenith=zenith0,fixedx=fixedx)
```

The proposal distributions are defined
```{r}
proposal.x <- mvnorm(S=diag(c(0.005,0.005)),n=nlocation(x0))
proposal.z <- mvnorm(S=diag(c(0.005,0.005)),n=nlocation(z0))
```
and all of the non-fixed locations are required to meet the conditions of
the mask
```{r,fig.show='hold'}
fit <- estelleMetropolis(model,proposal.x,proposal.z,iters=1000,thin=60,chains=1)
## Residuals are positive
sapply(fit$x,function(x) all(fixedx | model$residuals(chainLast(x)) > 0))
opar <- par(mfrow=c(2,1),mar=c(3,4,1,1)+0.1)
## x locations are at sea
plot(log.prior.x(chainLast(fit$x[[1]])),pch=16,cex=0.7,
     xlab="",ylab="Log prior x")
## z locations are at sea
plot(log.prior.z(chainLast(fit$z[[1]])),pch=16,cex=0.7,
     xlab="",ylab="Log prior z")
par(opar)
```

## Tuning

Once the chain meets the constraints of the model, the proposal
distributions are tuned.

The model is redefined
```{r}
x0 <- chainLast(fit$x)
z0 <- chainLast(fit$z)
model <- thresholdModel(twl$Twilight,twl$Rise,
                         twilight.model="LogNormal",
                         alpha=alpha,beta=beta,
                         logp.x=log.prior.x,logp.z=log.prior.z,
                         x0=x0,z0=z0,zenith=zenith0,fixedx=fixedx)
proposal.x <- mvnorm(S=diag(c(0.005,0.005)),n=nlocation(x0))
proposal.z <- mvnorm(S=diag(c(0.005,0.005)),n=nlocation(z0))
```
and a number of short runs are conducted
```{r}
fit <- estelleMetropolis(model,proposal.x,proposal.z,
                          iters=1000,thin=60,chains=2)
for(k in 1:3) {
  proposal.x <- mvnorm(chainCov(fit$x),s=0.2)
  proposal.z <- mvnorm(chainCov(fit$z),s=0.2)
  fit <- estelleMetropolis(model,proposal.x,proposal.z,
                            x0=chainLast(fit$x),
                            z0=chainLast(fit$z),
                            iters=1000,thin=60,chains=2)
}
```
The samples drawn are checked
```{r,fig.show='hold'}
opar <- par(mfrow=c(2,2),mar=c(3,5,2,1)+0.1)
for(k in 1:length(fit$x)) {
    matplot(t(fit$x[[k]][!fixedx,1,]),type="l",lty=1,col=ci.col,ylab="Lon")
    matplot(t(fit$x[[k]][!fixedx,2,]),type="l",lty=1,col=mn.col,ylab="Lat")
}
par(opar)
```

## Final Run

Once the proposals are tuned, a larger final sample is drawn
```{r}
proposal.x <- mvnorm(chainCov(fit$x),s=0.25)
proposal.z <- mvnorm(chainCov(fit$z),s=0.25)
fit <- estelleMetropolis(model,proposal.x,proposal.z,
                          x0=chainLast(fit$x),
                          z0=chainLast(fit$z),
                          iters=5000,thin=60,chains=4)
```

Corrections are made for fixed z
```{r}
## Fix the z locations.
fixedz <- which(fixed[-length(fixed)] > 0 & fixed[-length(fixed)]==fixed[-1])
for(k in 1:length(fit$z)) {
    fit$z[[k]][fixedz,1,] <- fixed.locations[fixed[fixedz],1]
    fit$z[[k]][fixedz,2,] <- fixed.locations[fixed[fixedz],2]
}
```


## Diagnostics


### Comparing Chains

The results from the parallel chains are compared for consistency.


A separate summary is generated for each chain
```{r}
s <- locationSummary(fit$x,time=model$time,collapse=F)
```


For each chain, plot the differences between the posterior mean of the
chain and the posterior means of the other chains, together with the
95% credible intervals for the deviation.
```{r,fig.show='hold'}
opar <- par(mfrow=c(2,2),mar=c(3,5,2,1)+0.1)
for(k in 1:length(s)) {
  ci <- as.matrix(s[[k]][,c("Lon.2.5%","Lon.97.5%")]-s[[k]][,"Lon.mean"])
  ci[fixedx,] <- NA
  matplot(s[[k]]$Time,ci,type="l",lty=1,ylab="Lon",xaxt="n",col=ci.col,
          cex.axis=0.7)
  axis.POSIXct(1,at=seq(round(min(s[[k]]$Time),"day"),
                   round(max(s[[k]]$Time),"day"),"months"),
               format="%d %b",cex.axis=0.7,las=2)
  for(j in 1:length(s))
    lines(s[[j]]$Time,s[[j]][,"Lon.mean"]-s[[k]][,"Lon.mean"],col=grp.pal[j])
}
par(opar)
```


```{r,fig.show='hold'}
opar <- par(mfrow=c(2,2),mar=c(3,5,2,1)+0.1)
for(k in 1:length(s)) {
  ci <- as.matrix(s[[k]][,c("Lat.2.5%","Lat.97.5%")]-s[[k]][,"Lat.mean"])
  ci[fixedx,] <- NA
  matplot(s[[k]]$Time,ci,type="l",lty=1,ylab="Lat",xaxt="n",col=ci.col,
          cex.axis=0.7)
  axis.POSIXct(1,at=seq(round(min(s[[k]]$Time),"day"),
                   round(max(s[[k]]$Time),"day"),"months"),
               format="%d %b",cex.axis=0.7,las=2)
  for(j in 1:length(s))
    lines(s[[j]]$Time,s[[j]][,"Lat.mean"]-s[[k]][,"Lat.mean"],col=grp.pal[j])
}
par(opar)
```


Plot the sequence of posterior mean locations from each chain.
```{r}
library(maptools)
data(wrld_simpl)
plot(wrld_simpl,xlim=xlim,ylim=ylim,col=map1.col,border=map2.col)
plot(elide(wrld_simpl,shift=c(360,0)),xlim=xlim,ylim=ylim,add=T,col=map1.col,
     border=map2.col)
for(k in 1:length(s)) {
  lines(s[[k]][,"Lon.mean"],s[[k]][,"Lat.mean"],
        col=rgb(t(col2rgb(grp.pal[k]))/255,alpha=0.4))
}
box()
```


Plot the posterior distribution of time spent at location derived from
each chain.
```{r,fig.show='hold'}
library(maptools)
data(wrld_simpl)
fixedz <- fixed[-length(fixed)] > 0 & fixed[-length(fixed)]==fixed[-1]
dt <- ifelse(fixedz,0,model$dt)
im <- locationImage(fit$z,xlim=xlim,ylim=ylim,nx=4*diff(xlim),ny=4*diff(ylim),
                    weight=dt,collapse=FALSE)
opar <- par(mfrow=c(2,2),mar=c(2,2,2,2)+0.1)
for(k in 1:length(im)) {
  image(im[[k]]$x,im[[k]]$y,im[[k]]$W,xlab="",ylab="",cex.axis=0.7, xlim=xlim0,
        ylim=ylim0)
  plot(wrld_simpl,add=T,col=map1.col,border=map2.col)
  plot(elide(wrld_simpl,shift=c(360,0)),add=T,col=map1.col,border=map2.col)
  points(fixed.locations, pch=16, cex=1)
  box()
}
par(opar)
```



### Diagnostic Predictions


The distribution of the posterior predictions of the twilight
errors for sunrise and sunset are compared to the  assumed
distribution of twilight errors
```{r}
pr.res <- chainApply(fit$x,model$residuals)
rs <- seq(0,quantile(pr.res,prob=0.98),length=200)
ys <- dlnorm(rs,alpha[1],alpha[2])
d1 <- density(pr.res[!fixedx & twl$Rise,])
d2 <- density(pr.res[!fixedx & !twl$Rise,])
plot(rs,ys,ylim=c(0,max(d1$y,d2$y,ys)),type="l",xlab="Minutes",ylab="",
     col=asm.col,lwd=2)
lines(d1,col=sr.col)
lines(d2,col=ss.col)
```

We plot 95% credible intervals of the twilight residuals as a ribbon
```{r,fig.show='hold'}
opar <- par(mar=c(4,5,2,1)+0.1)
qs <- t(apply(pr.res,1,quantile,prob=c(0.025,0.975)))
plot(model$time,qs[,2],type="n",col=ci.col,
     ylab="Minutes",xlab="",xaxt="n",cex.axis=0.7)
axis.POSIXct(1,at=seq(round(min(model$time),"day"),
                      round(max(model$time),"day"),"months"),
             format="%d %b",las=2,cex.axis=0.7)
polygon(c(model$time,rev(model$time)),
        c(qs[,2],rev(qs[,1])),col=ci.col,border=NA)
par(opar)
```

The posterior mean twilight times are shown against the light image.
```{r}
mn.twl <- .POSIXct(rowMeans(twl$Twilight+ifelse(twl$Rise,-60,60)*pr.res),"GMT")
d.lig <- readLig(paste0("GLS_files/", tag, "/", tag, "_000.lig"))
d.lig <- subset(d.lig,Date >= min(twl$Twilight)-72*60*60 & Date <= 
                  max(twl$Twilight)+72*60*60)
lightImage(d.lig,offset=offset)
tsimageLines(mn.twl[twl$Rise],offset=offset,col=sr.col,lwd=2)
tsimageLines(mn.twl[!twl$Rise],offset=offset,col=ss.col,lwd=2)
```

The posterior distribution of speeds is compared to the assumed
distribution of speeds
```{r}
spd <- function(x,z)
  pmax.int(trackDist2(x,z), 1e-06)/diff(as.numeric(twl$Twilight)/3600)
fixedz <- fixed[-length(fixed)] > 0 & fixed[-length(fixed)]==fixed[-1]
pr.spd <- chainApply2(fit$x,fit$z,spd)
ss <- seq(0,quantile(pr.spd,prob=0.98),length=200)
ys <- dgamma(ss,beta[1],beta[2])
ds <- density(pr.spd[!fixedz,])
plot(ss,ys,ylim=c(0,max(ds$y,ys[-1])),type="l",xlab="Km/h",ylab="",
     col=asm.col,lwd=2)
lines(ds,col=mn.col)
c(mean(pr.spd),sd(pr.spd))
```

The posterior distribution of SST is compared to the observed SST.
```{r}
reynolds <- sstTable(model$time,xlim=xlim,ylim=ylim,land=F,
                      src="Reynolds",path="SGAT")
pr.sst <- chainApply(fit$x,reynolds$sst)
qs <- t(apply(pr.sst,1,quantile,prob=c(0.025,0.975), na.rm=TRUE))
plot(model$time,qs[,2],ylim=c(0,max(qs)),type="n",
     ylab="SST",xlab="",xaxt="n",cex.axis=0.7)
polygon(c(model$time,rev(model$time)),c(qs[,2],rev(qs[,1])),
        col=ci.col,border=NA)
axis.POSIXct(1,at=seq(round(min(model$time),"day"),
                   round(max(model$time),"day"),"months"),
             format="%d %b",las=2,cex.axis=0.7)
points(twl$Twilight,twl$SST,pch=16,cex=0.5,col=mn.col)
```


## Summary

The optimal estimates are obtained by combining results from the four chains.
```{r,fig.show='hold'}
opar <- par(mfrow=c(2,1),mar=c(4,5,2,1)+0.1)
s <- locationSummary(fit$x,time=model$time)
matplot(s$Time,s[,c("Lon.mean","Lon.2.5%","Lon.97.5%")],
        type="l",lty=1,ylab="Lon",xaxt="n",xlab="",cex.axis=0.7,
        col=c(mn.col,ci.col,ci.col))
axis.POSIXct(1,at=seq(round(min(s$Time),"day"),
                      round(max(s$Time),"day"),"months"),
             format="%d %b",las=2,cex.axis=0.7)
matplot(s$Time,s[,c("Lat.mean","Lat.2.5%","Lat.97.5%")],
        type="l",lty=1,ylab="Lat",xaxt="n",xlab="",cex.axis=0.7,
        col=c(mn.col,ci.col,ci.col))
axis.POSIXct(1,at=seq(round(min(s$Time),"day"),
                      round(max(s$Time),"day"),"months"),
             format="%d %b",las=2,cex.axis=0.7)
par(opar)
```

More typically we are interested in projecting the samples onto a
map. The following function weights and bins samples to form weighted,
two dimensional frequency table.  The `locationImage` function can be
used to compute the posterior distribution of time spent at a location
by binning the z locations weighted by the time intervals between
twilights. To prevent the fixed locations from dominating the image
(as they are so localized), the fixed z locations are weighted out
```{r}
library(maptools)
data(wrld_simpl)
fixedz <- fixed[-length(fixed)] > 0 & fixed[-length(fixed)]==fixed[-1]
dt <- ifelse(fixedz,0,diff(as.numeric(model$time)/3600))
im <- locationImage(fit$z,xlim=xlim,ylim=ylim,nx=4*diff(xlim),ny=4*diff(ylim),
                    weight=dt)
image(im$x,im$y,im$W,xlab="Lon",ylab="Lat",cex.axis=0.7, xlim=xlim0,ylim=ylim0)
plot(wrld_simpl,add=T,col=map1.col,border=map2.col)
plot(elide(wrld_simpl,shift=c(360,0)),add=T,col=map1.col,border=map2.col)
points(fixed.locations, pch=16, cex=1)
box()
```

We can plot mean positions along the track
```{r}
xm <- locationMean(fit$x)
plot(xm,type="n",xlab="",ylab="",cex.axis=0.7, xlim=xlim0,ylim=ylim0)
plot(wrld_simpl,add=T,col=map1.col,border=map2.col)
plot(elide(wrld_simpl,shift=c(360,0)),add=T,col=map1.col,border=map2.col)
lines(xm,col=rgb(t(col2rgb(trk.col))/255,alpha=0.2))
points(xm,pch=16,cex=0.5,col=rgb(t(col2rgb(trk.col))/255,alpha=0.2))
points(fixed.locations, pch=16, cex=1)
box()
```

The track can be sliced into convenient intervals and each shown in turn
```{r}
r <- raster(nrows=4*diff(ylim0),ncols=4*diff(xlim0),xmn=xlim0[1],xmx=xlim0[2],
            ymn=ylim0[1],ymx=ylim0[2])
s <- slices(type="intermediate",breaks="month",mcmc=fit,grid=r)
for(k in sliceIndices(s)) {
  tm <- sliceInterval(s,k)
  sk <- slice(s,k)
  plot(sk,useRaster=F,main=sprintf("%s - %s",as.Date(tm[1]),as.Date(tm[2])),
       cex.axis=0.7, xlim=xlim0,ylim=ylim0)
  plot(wrld_simpl,add=T,col=map1.col,border=map2.col)
  plot(elide(wrld_simpl,shift=c(360,0)),add=T,col=map1.col,border=map2.col)
  points(fixed.locations, pch=16, cex=1)
  box()
}
```


```{r}
colnames(xm) <- c("Lon","Lat")
write.csv(cbind.data.frame(time=twl$Twilight,xm),
          paste0("GLS_files/", tag, "/", tagdata,"_1estelleSST.csv"),
          row.names=FALSE)
save(fit,file=paste0("GLS_files/", tag, "/", tagdata,"_fitSST.RData"))
##Saving Estelle output
write.csv(locationSummary(fit$x,time=fit$model$time),
          paste0("GLS_files/", tag, "/", tagdata,"_2estelleSST.csv"),
          row.names=FALSE)
toc()
```

<!--
```{r VersionInfo,results='asis',echo=FALSE}
cat(R.version.string,"\n",
    "SGAT: ",as.character(packageVersion("SGAT")),"\n",
    "BAStag: ",as.character(packageVersion("BAStag")),sep="")
```
-->
